---
title: "Exploratory Data Analyst - Project 2"
author: "Mehdi HAMDOUNE"
date: "3/10/2021"
output:
  pdf_document: default
  html_document: default
---
# **Exploratory Data Analyst - Project 2**

## *Instroduction*

Fine particulate matter (PM2.5) is an ambient air pollutant for which there is strong evidence that it is harmful to human health. In the United States, the Environmental Protection Agency (EPA) is tasked with setting national ambient air quality standards for fine PM and for tracking the emissions of this pollutant into the atmosphere. Approximatly every 3 years, the EPA releases its database on emissions of PM2.5. This database is known as the National Emissions Inventory (NEI). You can read more information about the NEI at the EPA National Emissions Inventory web site.

For each year and for each type of PM source, the NEI records how many tons of PM2.5 were emitted from that source over the course of the entire year. The data that you will use for this assignment are for 1999, 2002, 2005, and 2008.

PM2.5 Emissions Data contains all of the PM2.5 emissions data for 1999, 2002, 2005, and 2008. For each year, the table contains number of tons of PM2.5 emitted from a specific type of source for the entire year.

Source Classification Code Table provides a mapping from the SCC digit strings in the Emissions table to the actual name of the PM2.5 source. The sources are categorized in a few different ways from more general to more specific and you may choose to explore whatever categories you think are most useful. For example, source “10100101” is known as “Ext Comb /Electric Gen /Anthracite Coal /Pulverized Coal”.

The overall goal of this project is to explore the National Emissions Inventory database and see what it say about fine particulate matter pollution in the United states over the 10-year period 1999–2008. To do so, we're going to answer these answers :

1 - Have total emissions from PM2.5 decreased in the United States from 1999 to 2008?

2 - Have total emissions from PM2.5 decreased in the Baltimore City, Maryland from 1999 to 2008?

3 - Which of the four sources have seen decreases in emissions from 1999–2008 for Baltimore City? Which have seen increases in emissions from 1999–2008?

4 - Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

5 - How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?

6 - Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California . Which city has seen greater changes over time in motor vehicle emissions?

## *Importing libs and Dataset *

```
# Importing The LIbrairies
library(tidyr)

# Importing The Dataset
Dataset <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```
```{r, echo = FALSE,  warning = FALSE, message = FALSE } 
setwd("D:/Auto_formation/Coursera_Exploratory_Data_Analysis/Project 2")

# Importing The LIbrairies
library(tidyr)
library(ggplot2)

# Importing The Dataset
Dataset <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
head(Dataset, 5)
```


## *Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? *

```
#Calculating the total emission for each year
totalEm <- aggregate(Emissions ~ year, Dataset, sum)

#plotting the total emission for each year
plot(totalEm$year, totalEm$Emissions, type = 'o', col = "blue", main = "Total US PM2.5 Emissions by Year", 
     ylab = expression("Total US PM2.5 Emissions"), xlab = "Year")

```
```{r, echo = FALSE,  warning = FALSE, message = FALSE } 
#Calculating the total emission for each year
totalEm <- aggregate(Emissions ~ year, Dataset, sum)

#plotting the total emission for each year
plot(totalEm$year, totalEm$Emissions, type = 'o', col = "blue", main = "Total US PM2.5 Emissions by Year", 
     ylab = expression("Total US PM2.5 Emissions"), xlab = "Year")

```


## *Have total emissions from PM2.5 decreased in the Baltimore City from 1999 to 2008? *

```
# Subsetting The baltimore City's Emissions
Em_Baltimore <- subset(Dataset, fips == "24510")

#Calculating the total emission for each year
totalEm_bal <- aggregate(Emissions ~ year, Em_Baltimore, sum)

#plotting the total emission for each year
plot(totalEm_bal$year, totalEm_bal$Emissions, type = 'o', col = "blue", main = "Total PM2.5 Emissions by Year in   Baltimore City", ylab = expression("Total PM2.5 Emissions in Baltimore City"), xlab = "Year")
```
```{r, echo = FALSE,  warning = FALSE, message = FALSE } 
# Subsetting The baltimore City's Emissions
Em_Baltimore <- subset(Dataset, fips == "24510")

#Calculating the total emission for each year
totalEm_bal <- aggregate(Emissions ~ year, Em_Baltimore, sum)

#plotting the total emission for each year
plot(totalEm_bal$year, totalEm_bal$Emissions, type = 'o', col = "blue", main = "Total PM2.5 Emissions by Year in    Baltimore City", ylab = expression("Total PM2.5 Emissions in Baltimore City"), xlab = "Year")
```


## *Which of The four sources have seen decreases in emissions from 1999–2008 for Baltimore City? Which have seen increases in emissions from 1999–2008?*

```
# Subsetting The baltimore City's Emissions
Em_Baltimore <- subset(Dataset, fips == "24510")

#Calculating the total emission for each year and for each type
totalEm_bal_type <- aggregate(Emissions ~ year + type, Em_Baltimore, sum)

#Plotting The Baltimore emission by year and type
ggplot(totalEm_bal_type, aes(year, Emissions, col = type)) +
  geom_line() +
  geom_point() +
  ggtitle("Total PM2.5 Emissions by Type and Year in Baltimore City") +
  ylab(expression("Total PM2.5 Emissions in Baltimore City")) +
  xlab("Year") +
  scale_colour_discrete(name = "Type of sources")+
  theme(plot.title = element_text(hjust = 0.5))
```
```{r, echo = FALSE,  warning = FALSE, message = FALSE } 
# Subsetting The baltimore City's Emissions
Em_Baltimore <- subset(Dataset, fips == "24510")

#Calculating the total emission for each year and for each type
totalEm_bal_type <- aggregate(Emissions ~ year + type, Em_Baltimore, sum)

#Plotting The Baltimore emission by year and type
ggplot(totalEm_bal_type, aes(year, Emissions, col = type)) +
  geom_line() +
  geom_point() +
  ggtitle("Total PM2.5 Emissions by Type and Year in Baltimore City") +
  ylab(expression("Total PM2.5 Emissions in Baltimore City")) +
  xlab("Year") +
  scale_colour_discrete(name = "Type of sources")+
  theme(plot.title = element_text(hjust = 0.5))
```


## *Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?*

```
# Subsetting The Emissions from coal combustion-related in the US 
SCC_coal <- SCC[grepl("coal", SCC$Short.Name, ignore.case = T),]
Em_coal <- Dataset[Dataset$SCC %in% SCC_coal$SCC,]

#Calculating the total emission for each year and for each type
totalEm_col <- aggregate(Emissions ~ year, Em_coal, sum)

#Plotting The Baltimore emission by year and type
ggplot(totalEm_col, aes(year, Emissions)) +
  geom_line() +
  geom_point() +
  ggtitle("Total US PM2.5 Emissions from coal combustion-related by year") +
  ylab(expression("Total US PM2.5 Emissions from coal combustion-related")) +
  xlab("Year") +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r, echo = FALSE,  warning = FALSE, message = FALSE } 
# Subsetting The Emissions from coal combustion-related in the US 
SCC_coal <- SCC[grepl("coal", SCC$Short.Name, ignore.case = T),]
Em_coal <- Dataset[Dataset$SCC %in% SCC_coal$SCC,]

#Calculating the total emission for each year and for each type
totalEm_col <- aggregate(Emissions ~ year, Em_coal, sum)

#Plotting The Baltimore emission by year and type
ggplot(totalEm_col, aes(year, Emissions)) +
  geom_line() +
  geom_point() +
  ggtitle("Total US PM2.5 Emissions from coal combustion-related by year") +
  ylab(expression("Total US PM2.5 Emissions from coal combustion-related")) +
  xlab("Year") +
  theme(plot.title = element_text(hjust = 0.5))
```


## *How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?*

```
# Subsetting The baltimore City's Emissions from motor vehicule
Em_Baltimore_Mtr <- subset(Dataset, fips == "24510" & type == "ON-ROAD")

#Calculating the total emission for each year and for each type
totalEm_bal_Mtr <- aggregate(Emissions ~ year, Em_Baltimore_Mtr, sum)

#Plotting The Baltimore emission by year and type
ggplot(totalEm_bal_Mtr, aes(year, Emissions)) +
  geom_line() +
  geom_point() +
  ggtitle("Total PM2.5 Emissions by Year in Baltimore City from motor vehicule") +
  ylab(expression("Total PM2.5 Emissions in Baltimore City from motor vehicule")) +
  xlab("Year") +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r, echo = FALSE,  warning = FALSE, message = FALSE } 
# Subsetting The baltimore City's Emissions from motor vehicule
Em_Baltimore_Mtr <- subset(Dataset, fips == "24510" & type == "ON-ROAD")

#Calculating the total emission for each year and for each type
totalEm_bal_Mtr <- aggregate(Emissions ~ year, Em_Baltimore_Mtr, sum)

#Plotting The Baltimore emission by year and type
ggplot(totalEm_bal_Mtr, aes(year, Emissions)) +
  geom_line() +
  geom_point() +
  ggtitle("Total PM2.5 Emissions by Year in Baltimore City from motor vehicule") +
  ylab(expression("Total PM2.5 Emissions in Baltimore City from motor vehicule")) +
  xlab("Year") +
  theme(plot.title = element_text(hjust = 0.5))
```


## *Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California. Which city has seen greater changes over time in motor vehicle emissions*

```
# Subsetting The baltimore City and Los Angeles Emissions from motor vehicle
Em_Mtr <- subset(Dataset,  type == "ON-ROAD" & fips %in% c("24510","06037"))

#Calculating the total emission for each year and for each city
totalEm_Mtr <- aggregate(Emissions ~ year + fips, Em_Mtr, sum)

#Plotting The total emission by year and city
ggplot(totalEm_Mtr, aes(year, Emissions, col = fips)) +
  geom_line() +
  geom_point() +
  ggtitle("Total PM2.5 Emissions by Year and by City from motor vehicule") +
  ylab(expression("Total PM2.5 Emissions from motor vehicule")) +
  xlab("Year") +
  scale_colour_discrete(name = "City", labels = c("Los Angeles", "Baltimore"))+
  theme(plot.title = element_text(hjust = 0.5))
```
```{r, echo = FALSE,  warning = FALSE, message = FALSE } 
# Subsetting The baltimore City and Los Angeles Emissions from motor vehicle
Em_Mtr <- subset(Dataset,  type == "ON-ROAD" & fips %in% c("24510","06037"))

#Calculating the total emission for each year and for each city
totalEm_Mtr <- aggregate(Emissions ~ year + fips, Em_Mtr, sum)

#Plotting The total emission by year and city
ggplot(totalEm_Mtr, aes(year, Emissions, col = fips)) +
  geom_line() +
  geom_point() +
  ggtitle("Total PM2.5 Emissions by Year and by City from motor vehicule") +
  ylab(expression("Total PM2.5 Emissions from motor vehicule")) +
  xlab("Year") +
  scale_colour_discrete(name = "City", labels = c("Los Angeles", "Baltimore"))+
  theme(plot.title = element_text(hjust = 0.5))
```